Humann2 bar plot by myself

### 2019-10-17 scripts/08_barplot_species_pathway.Rmd
that stupid humann2 bar plot script finally worked

```{r}
# parse the distinct pathway feature in the lefse results from the class: CR and tox
pw <- read_csv('../data/unique_enriched_pathways.csv') %>% 
  # something has a f_ at the beginning need to trim that off too
  mutate(pathway = str_replace(pathway, '^f_','')) %>% 
  # extract the pathway id from the whole messy string
  mutate(pwid = str_extract(pathway, '^.+_PWY|^PWY\\d*_\\d{3,4}')) %>% 
  mutate(pwid = str_replace_all(pwid, '_','-'))

```

## Make species level bar plot

```{r}
# check one last time whether the damn sampleids align
cnts <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified.tsv') %>% 
  rename_all(funs(str_replace(., '_Abundance$','')))

# omg, its really hard to sort the pheno table the way I wanted, I will have to resort the cnts table to match
# reorder the relab column to be the same with pheno table
cnts_or <- bind_cols(cnts[,1],cnts[,pheno$Sampleid])

# just to be sure
all.equal(colnames(cnts_or)[2:ncol(cnts_or)], pheno$Sampleid)

# write out the reordered cnts file
cnts_or %>% 
  write_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv' )

# write out the pheno as metadata
pheno %>% 
  select(Toxicity) %>% 
  t() %>% 
  write.table(str_glue('../data/tox_pheno.tsv'), sep = '\t', quote = F, row.names = T, col.names = F)

cnts_or_ <- cnts_or %>% 
  column_to_rownames('# Pathway')

in_meta_t <- data.frame(t(pheno %>% 
  select(Toxicity)), stringsAsFactors = FALSE)

names(in_meta_t) <- names(cnts_or_)
cat_pw_table <- rbind(in_meta_t, cnts_or_ )

cat_pw_table %>% 
  write.table('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.pcl',
              quote = F,
              col.names=FALSE, 
              row.names=TRUE, 
              sep="\t")

```

```{bash}
# guess we need to cat the pheno and cnts together?
cat ../data/tox_pheno.tsv \
    ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv > \
    ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.pcl
```

```{bash}
# the actual cmd to plot 
humann2_barplot  \
  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.pcl \
  --focal-feature PWY-4242 \
  --sort similarity metadata \
  --focal-metadatum Toxicity --last-metadatum Toxicity \
  --scaling pseudolog \
  --dimensions 10 5 \
  -o ../figs/PWY-4242_pw.png
```

```{bash}
# without metadata
humann2_barplot  \
  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv \
  --focal-feature PWY-4242 \
  --sort sum \
  --dimensions 10 5 \
  -o ../figs/PWY-4242_pw.png

  #   --scaling pseudolog \
```

```{bash}
# the actual cmd to plot 
`humann2_barplot  \
  --input ../data/humann2_res/test_ARO-PWY_pw_abundance.tsv \
  --focal-feature ARO-PWY \
  --sort metadata \
  --focal-metadatum Toxicity --last-metadatum Toxicity \
  --scaling pseudolog \
  --dimensions 10 5 \
  -o ../figs/ARO-PWY_pw.png`

```

>>>>

## Make species level bar plot using humann2 script

### Create the correct pheno and relab table format

```{r}
# ok so the correct format should be header at the first row, and phenotype and then the abundance values 
# pheno table 
read_csv('../data/cleaned_pheno/current_pheno.csv') %>% 
  filter(Sampleid %in% valid_s$X1) %>% 
  select(-Censor_YN, -Censor_reason) %>% 
  # USE THE Toxicity PHENO FIRST
  select(Sampleid, Toxicity) %>% 
  t() %>% 
  write.table('../data/horizontal_pheno/tox_wi_sample_header.txt', quote = F, row.names = T, col.names = F, sep = '\t')

# reordered relab table but without header
read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv') %>% 
  write_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_wo_header.tsv', col_names = F)
```

### row bind them together

```{bash}
# guess we need to cat the pheno and cnts together?
cat ../data/horizontal_pheno/tox_wi_sample_header.txt \
    ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_wo_header.tsv > \
    ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv
```

FINALLY WORKED!!!!!!!!!!!!!!!!!!!!!!

```{bash}
# the actual cmd to plot 
humann2_barplot  \
  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv \
  --focal-feature PWY-4242 \
  --sort similarity sum metadata \
  --focal-metadatum Toxicity --last-metadatum Toxicity \
  --scaling pseudolog \
  --dimensions 10 5 \
  -o ../figs/PWY-4242_pw.png
```

### 2019-10-17 scripts/09_barplot_myself.Rmd
that stupid humann2 bar plot script finally worked

sig_num <- siglefse %>% 
  split(f = .$pwid) %>% 
  map_dfc(~ nrow(.)) %>% 
  gather(key = 'pw', value = 'taxa_num') %>% 
  arrange(desc(taxa_num)) %>% 
  separate(pw, into = c('pwid','pw'), sep = ': ') %>% 
  # for plotting reason filer on >= 9 rows
  filter(taxa_num >= 9)

@@

```{r}
species_compo_barplot <- function(PHENOVAR, PATHWAYID) {
  # TWO INPUTS: 
  # The pheno variable: CR, Toxicity, ORR
  # The full PATHWAYID
  # get the pheno lable of the samples

@@

  # the ggplot
  ggplot() +
    geom_bar(data = final_cnt, 
             aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
    geom_rect(data = background_col, 
              aes(xmin = xstart, xmax = xend, ymin = -Inf, ymax = Inf), 
              fill = background_col$col, 
              alpha = 0.2) +
    geom_bar(data = final_cnt, aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
    scale_fill_manual(values = hex ) +
    labs(y = 'relative abundance',
         x = 'Sample',
         title = str_glue('Response to Toxicity NO and YES:\nspecies composition of the pathway abundance\n{PATHWAYID}')) +
    theme_light() +
    theme(
      axis.text.x =  element_text(angle = 40, hjust = 1, size = 8),
      panel.grid.major = element_blank(), 
      axis.ticks.x = element_blank(),  
      panel.grid.minor = element_blank()
      ) +
    ggsave(str_glue('../figs/species_pw/{PATHWAYID}.jpg'), width = 12, height = 8, dpi = 300)
}

sig_num$pwid %>% 
  map(~  species_compo_barplot('Toxicity', .))

>>>>

sig_num <- siglefse %>% 
  split(f = .$pwid) %>% 
  map_dfc(~ nrow(.)) %>% 
  gather(key = 'pw', value = 'taxa_num') %>% 
  arrange(desc(taxa_num)) %>% 
  separate(pw, into = c('pwid','pw'), sep = ': ', remove = F) %>% 
  # for plotting reason filer on >= 9 rows
  filter(taxa_num >= 9)

@@

```{r}
species_compo_barplot <- function(PATHWAYID, PWFULLNAME) {
  # TWO INPUTS: 
  # The pheno variable: CR, Toxicity, ORR
  # The full PATHWAYID
  # get the pheno lable of the samples

@@

  # the ggplot
  ggplot() +
    geom_bar(data = final_cnt, 
             aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
    geom_rect(data = background_col, 
              aes(xmin = xstart, xmax = xend, ymin = -Inf, ymax = Inf), 
              fill = background_col$col, 
              alpha = 0.2) +
    geom_bar(data = final_cnt, aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
    scale_fill_manual(values = hex ) +
    labs(y = 'relative abundance',
         x = 'Sample',
         title = str_glue('Response to Toxicity NO (blue background) and YES (red background):\nspecies composition of the pathway abundance\n{PATHWAYID}: {PWFULLNAME}')) +
    theme_light() +
    theme(
      axis.text.x =  element_text(angle = 40, hjust = 1, size = 8),
      panel.grid.major = element_blank(), 
      axis.ticks.x = element_blank(),  
      panel.grid.minor = element_blank()
      ) +
    ggsave(str_glue('../figs/species_pw/{PATHWAYID}.jpg'), width = 12, height = 8, dpi = 300)
}

map2(sig_num$pwid, sig_num$pw, ~ species_compo_barplot(.x, .y))

sig_num$pw
```
